---
title: "Where are the South African Expats ?"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)

doublePie <- function(location.dt) {
  ####################################################################################################
  # location.dt is a data table and needs to contain 3 columns - 
  #     continent_map lists all the continents (group), unique, max 8 colours assigned
  #     country_map   lists all the countries in the continent (sub-group), unique, top 10 labeled
  #     count         counts per country (sub-group)
  # Result a pie/donut map of the groups and sub-group
  # Uses the package: plotrix, maps one pie chart on top of another
  ####################################################################################################
  
  # PiDonut using floating
  library(plotrix)
  
  # aggregate data for the continent pie chart
  continent_data.dt <- location.dt[,sum(count), by = "continent_map" ] #automatically order by continent_map
  # colour blind friendly colours
  continent_colours <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
  setDT(continent_data.dt)[,colours := continent_colours[1:nrow(continent_data.dt)]]
  continent_data.dt <- continent_data.dt[order(continent_map),]
  
  # format labels to display country and count
  country_data.dt <- location.dt[order(-count),]
  setDT(country_data.dt)[,country_label := paste(country_data.dt$country_map, ": ", country_data.dt$count, sep = "")]
  if (nrow(country_data.dt)>10) {country_data.dt$country_label[11:nrow(country_data.dt)] ="" }#only display top 10
  #country_labels <- paste(country_data.dt$country_map, ": ", country_data.dt$count, sep = "")
  
  #add same colours to country data by doing a join on the continent field
  country_data.dt <- country_data.dt[continent_data.dt,on="continent_map"]
  # order country data by continent so it will line up with continent pie chart
  country_data.dt <- country_data.dt[order(continent_map),]
  
  # coordinates for the center of the chart
  center_x <- 0.5
  center_y <- 0.5
  
  plot.new()
  
  # draw country pie chart first
  country_chart <-
    floating.pie(
      xpos = center_x,
      ypos = center_y,
      x = country_data.dt$count,
      radius = 0.35,
      border = "white",
      col = country_data.dt$colours
    )
  # add labels for country pie chart
  pie.labels(
    x = center_x,
    y = center_y,
    angles = country_chart,
    labels = country_data.dt$country_label,
    radius = 0.38,
    bg = NULL,
    cex = 0.8,
    font = 2,
    col = "gray40"
  )
  
  # overlay continent pie chart
  continent_chart <-
    floating.pie(
      xpos = center_x,
      ypos = center_y,
      x = continent_data.dt$V1,
      radius = 0.25,
      border = "white",
      col = continent_data.dt$colours
    )
  
  # add labels for continent pie chart
  pie.labels(
    x = center_x,
    y = center_y,
    angles = continent_chart,
    labels = continent_data.dt$continent_map,
    radius = 0.125,
    bg = NULL,
    cex = 0.8,
    font = 2,
    col = "white"
  )

}
################################################################################################
#  find the orgins of the tweets and map them in a single and double pie graph
#  exclude tweets from Africa or where the continent is unknown
#  Call the two functions:
#     singlePie and doublePie
################################################################################################

  library(RMySQL)
  library(DBI)
  library(data.table)
  library(tidyselect)
  library(wordcloud)
  library(ggplot2)
  
########### Query Database
  #Connect with MySQL locally hosted database 'twitterfeed'
  user <- '####'
  password <- '#######'
  dbname <- '#######'
  host <- '########'
  drv <- dbDriver("MySQL")
  mydb <- dbConnect(drv, user=user, password=password, dbname=dbname, host=host)
  #Query database
  #what are all the locations mapped
  mappings.dt <- as.data.table(dbGetQuery(mydb,"
        select * from location_map
        "))
  #who sent all the tweets
  lean_tweets.dt <- as.data.table(dbGetQuery(mydb,"
        select clean_loc, user_id, created_at, word from tweet_data_lean 
        "))
  dbDisconnect(mydb)

########### Prepare data  
  #combine tweets with their mapped locations
  tweets.dt <- mappings.dt[lean_tweets.dt, on="clean_loc", nomatch=0]
  tweets.dt <- tweets.dt[!(is.na(tweets.dt$continent_map)),] #remove unmapped entries
  
  #find the first and last date for full date interval
  #- if single date set both to single date format "2019-10-21"
  setorder(tweets.dt,created_at) #order from earliest to latest date
  start_date <- tweets.dt[1]$created_at #earliest date
  end_date <- tweets.dt[nrow(tweets.dt)]$created_at #latest date
  tweets.dt <- tweets.dt[created_at>=start_date & created_at<=end_date] #Filter by start and end date

  ############# Location Plots
  ########### Prepare location data for location graphs
  #Unique user for the date interval selected - last tweet location seen as the truth
  twitters.dt <- unique(tweets.dt, by="user_id",fromLast=TRUE)
  # nonAfrican Pie charts
  # remove entries where continent is Unknown or continent is Africa
  NonAfrican.dt<- twitters.dt[!(twitters.dt$continent_map %in% c('Unknown','Africa')),] 
  # count words per country and continent
  loc_country.dt <- NonAfrican.dt[, .(count = .N), by = list(continent_map,country_map)]
  # count words per country and continent
  loc_continent.dt <- NonAfrican.dt[, .(count = .N), by = list(continent_map)]
  
  ########### Prepare location data - each word SA vs the world
  # remove entries where continent is Unknown
  loc_country_word.dt <- tweets.dt[!(continent_map=='Unknown')]
  loc_country_word.dt <- loc_country_word.dt[, .(count = .N), by = list(word, continent_map, country_map)]
  
  # Sum up the words used outside South Africa
  lean_words_country_counted_noZAR.dt <- loc_country_word.dt[!(country_map=='South Africa') ]
  lean_words_country_counted_noZAR.dt <- lean_words_country_counted_noZAR.dt[, .(count = sum(count)), by = list(word)]
  lean_words_country_counted_noZAR.dt <- lean_words_country_counted_noZAR.dt[, country_map := "Rest of the world"] 
  # Sum up the words used inside South Africa
  lean_words_country_counted_ZAR.dt <- loc_country_word.dt[(country_map=='South Africa') ]
  lean_words_country_counted_ZAR.dt <- lean_words_country_counted_ZAR.dt[,continent_map:=NULL]
  # Merge the two
  words_counted_ZAR_vs_theWorld.dt <- merge(lean_words_country_counted_ZAR.dt,lean_words_country_counted_noZAR.dt,all = TRUE)
  ########### Prepare word data
  # Frequency of words in tweets regardless of location
  words.dt <- tweets.dt[, .(count = .N), by = list(word)]
  # Words by continent - Unknown exlcuded
  words_by_continent.dt <- tweets.dt[!(continent_map=='Unknown')]
  words_by_continent.dt <- words_by_continent.dt[, .(count = .N), by = list(word,continent_map)]

```

## Aim to

* Determine the continent and countries where South African ex-pats are staying.
* Map the change in locations over time in monthly intervals.

## Methodology

In this proof of concept the origin of tweets containing specific South African words will be used to determine the location of South African expats.

* Harvest tweets sent containing specific South African slang words.  The slang words was originally sourced from the [Brand South Africa's website](https://www.brandsouthafrica.com/people-culture/culture/south-african-english).  The list is reduced to words uniquely South African in any context.  
The following 17 words are included:

```{r echo=TRUE, fig.height = 3,fig.width=3, fig.align="center"}
  #word cloud  
  wordcloud(words = words.dt$word, freq = words.dt$count, min.freq = 20,
            max.words=50, random.order=FALSE, rot.per=0.15,
            colors=brewer.pal(8, "RdYlGn"))
```

* Refer to the location field in the twitter record to determine the origin of each tweet.

* Map the continent and country of the origin of the tweets, excluding Africa.

## Test quality of results

* The majority of tweets per word need to come from South Africa else the word is not uniquely South African.

* In addition the resulting countries are similar to surveys conducted by the site [wheredidwego.com](https://www.wheredidwego.com/) listing the following 10 countries as the top destinations:
New Zealand, Australia, UK, Canada, USA, Netherlands, Ireland, Germany, UAE, Qatar

## Roadblock

The location field is a free text field.  Many users do not complete the field which means these tweets would need to be discarded.
Where the location field is completed, the entries varies. Each of these entries needs to be mapped to a valid country and continent.
The expectation is the need to map will decrease as more mapping is added.

After harvesting tweets for 14 days there was still around 300 locations that needed to be mapped each day which made the excercise unsustainable beyond a proof of concept.  If the proof of concept pass, the next step is to connect to the Google Maps API to allow a more efficient determination of the location origin of the a tweet.

## Resulting data

These are the resutls from the 14 day excercise - 12 October to 26 October 2019.  

All the words in scope are used more than 50% of the time by twitters from South Africa validating the words as typical South African.  

```{r echo=TRUE}
  # bar chart plot using ggplot2
   p1 <- ggplot(words_counted_ZAR_vs_theWorld.dt,aes(x=word,y=count,fill=factor(country_map)))+
    geom_bar(stat="identity",position="dodge")+
    scale_fill_discrete(name="Region",
                        breaks=c('South Africa', 'Rest of the world'),
                        labels=c("South Africa", "The rest of the world"))+
    xlab("Word")+ylab("Counted")+
    theme(plot.title = element_text(hjust = 0.5,size=14, face="bold"))+
    ggtitle(paste("Tweets sent between",start_date,"and",end_date))
  p1
```  

The only exception is the word Rooibos.  Although less than 50% of the tweets come from South Africa, Africa is still the largest single source of tweets containing the word Rooibos and Rooibos is still included as a typical South African word.

```{r echo=TRUE}
  # single pie chart - ggplot2
  s_word <- 'rooibos'
  w_count <- sum(words_by_continent.dt[word == s_word]$count)
  p3 <- ggplot(words_by_continent.dt[word == s_word], 
               aes(x="", y=count, group=continent_map, fill=continent_map)) +
   geom_bar(width = 1, stat = "identity") +
   coord_polar("y", start=0) +
   theme(axis.text = element_blank(),
         axis.ticks = element_blank(),
         panel.grid  = element_blank(),
         plot.title = element_text(hjust = 0.5,size=14, face="bold"))+
   ggtitle(paste(s_word,w_count,sep=","))+
   labs(fill = "Continent",x="", y="")
  p3
```  

The Excercise fail the test of likely countries to appear on the list. Refer to the list of countries and continents repesented in the double pie chart.

```{r echo=TRUE, fig.height = 12,fig.width=12, fig.align="center"}
doublePie(loc_country.dt)
```

The two most popular countries listed on the site [wheredidwego.com](https://www.wheredidwego.com/) for South Africans to migrate to are Australia and New Zealand.  Refering to the chart, Australia and New Zealand are dwarfed by the USA and UK in the twitter stats.

It is possible that countries have different cultures when it comes to sending tweets and South African expats in the USA and UK tweet alot more than their cousins in Australia and New Zealand.  The location origin of tweets can not be taken as a single indicator of where South African have settled.

## What next ?

It is interesting to note if Australia and New Zealand is taken out of the equation, the list of most populater countries are very similar.  This should be explored.
In addition if a more accurate and efficient determination of origin with the google maps API is used the result can be measured with more confidence.